---
title: "HTTP Server Log"
author: "Vasudeo Maurya"
date: "25 September 2018"
output: html_document
---

**Kint Opts**
```{r setup, include=FALSE}
# for development
knitr::opts_chunk$set(echo=TRUE, eval=TRUE, error=TRUE, warning=TRUE, message=TRUE, cache=FALSE, tidy=FALSE, fig.path='figures/')
# for production
#knitr::opts_chunk$set(echo=TRUE, eval=TRUE, error=FALSE, warning=FALSE, message=FALSE, cache=FALSE, tidy=FALSE, fig.path='figures/')
```
**Load Libs**   
```{r load_libs, echo=TRUE}
library(dplyr)
library(stringr)
library(ggplot2)
library(wordcloud)
library(tidyverse)
```
**Read Data**   
```{r read_data, echo=TRUE}
dfrDataFile <- read_log("./data/apache_dataset.log",skip =0,col_names = FALSE)
head(dfrDataFile)
```
**Rename Column**   
```{r rename_column, echo=TRUE}
df <- rename(dfrDataFile, RemoteHost=X1,Id=X2,HTTPUser=X3,DateTime=X4,RequestLine=X5,StatusCode=X6,Size=X7)
```

**Dataframe Stucture**   
```{r structure, echo=TRUE}
str(df)
```
**Splitting Date and Time**   
```{r split_date_time, echo=TRUE}
Hours <- format(strptime(df$DateTime,"%d/%B/%Y:%H:%M:%S") ,format = "%H:%M:%S")
Dates <- format(strptime(df$DateTime,"%d/%B/%Y:%H:%M:%S") ,format = "%d/%b/%Y")
#adding Dates to df as column
df$Dates <- Dates
#adding Hours to df as column
df$Hours <- Hours
#Below outputs the dates in the international standard format
df$Dates=as.Date(df$Dates,"%d/%b/%Y")
##Dropping DateTime Column after splitting date and time from df
df <- subset( df, select = -DateTime)
head(df)
```
**Graph showing statuscode**   
```{r graph_status, echo=TRUE}
ggplot(data=df, aes(x=format(StatusCode))) + geom_bar() + xlab("Status") + ylab("Count") + ggtitle("Status")
```
**Graph showing traffic to site**   
```{r graph_traffic_to_site, echo=TRUE}
reqs = as.data.frame(table(df$Dates))
ggplot(data = reqs, aes(x=as.Date(Var1), y=Freq)) + geom_point() + geom_line(colour = "blue") + xlab("Month") + ylab("Requests") + labs(title="Traffic to Site")
```
**Selecting data with successful statuscode**   
```{r status_200, echo=TRUE}
df_conn <- filter(df, StatusCode == "200")

#splitting date into day/month/year as separate column
df_conn <- data.frame(df_conn,
                    Year = as.numeric(format(df_conn$Dates, format = "%Y")),
                    Month = as.numeric(format(df_conn$Dates, format = "%m")),
                    Day = as.numeric(format(df_conn$Dates, format = "%d")))
head(df)
```
**Graph showing total hits per day**   
```{r hits_per_day, echo=TRUE}
thpd = as.data.frame(table(df_conn$Day,df_conn$Month))
thpd <- rename(thpd, Days=Var1,Months=Var2)
ggplot(data = thpd, aes(x=Days, y=Freq,group=Months,colour=Months)) + geom_point() + geom_line() + xlab("Day") + ylab("Freq") + labs(title="Total Hits Per Day")
```
**Graph showing Connection to Server Per Month**   
```{r connection_server, echo=TRUE}
dfcts <- data.frame(table(df_conn$RemoteHost,df_conn$Month))
dfcts.Sort <- arrange(dfcts,desc(Freq))
dfcts.Sort <- rename(dfcts.Sort,Host=Var1,Month=Var2)
head(dfcts.Sort)
ggplot(data = dfcts.Sort, aes(x=Month, y=Freq)) + geom_bar(stat = "identity") + xlab("Months") + ylab("Freq") + labs(title="Connection to Server Per Month")
```
**Page requested from server**   
```{r page_req, echo=TRUE}
dfreq <- data.frame(table(df_conn$RequestLine,df_conn$Month))
dfreq.Sort <- arrange(dfreq,desc(Freq)) 
dfreq.Sort <- rename(dfreq.Sort,Page=Var1,Month=Var2)
head(dfreq.Sort)
```
**Data downloaded by each individual host**   
```{r host_size, echo=TRUE}
dfSize <- aggregate(Size ~ RemoteHost + Month, data = df_conn, sum)
dfSize <- filter(dfSize, !grepl("0",dfSize$Size))
dfSize.Sort <- arrange(dfSize,desc(Size))
head(dfSize.Sort)
ggplot(data=dfSize.Sort, aes(x=format(Month))) + geom_bar() + xlab("Months") + ylab("Download Size") + ggtitle("Total Download Size Per Month")
```
**Download Size Per Day**   
```{r size_day, echo=TRUE}
dfSizepd <- aggregate(Size ~ Day + Month, data = df_conn, sum)
dfSizepd <- filter(dfSizepd, !grepl("0",dfSizepd$Size))
ggplot(data = dfSizepd, aes(x=Day, y=Size,group=Month,colour=Month)) + geom_point() + geom_line() + xlab("Day") + ylab("Download Size") + labs(title="Total Download Size Per Day")
```

**Data Sent Out Per Month**   
```{r size_month, echo=TRUE}
dfPageSize <- aggregate(Size ~ RequestLine + Month, data = df_conn, sum)
dfPageSize <- filter(dfPageSize, !grepl("0",dfPageSize$Size))
dfPageSize.Sort <- arrange(dfPageSize,desc(Size))
head(dfPageSize.Sort)
ggplot(data=dfPageSize.Sort, aes(x=format(Month))) + geom_bar() + xlab("Months") + ylab("Data Sent Out") + ggtitle("Total Data Sent Out Per Month")
```
**Which host has connected the maximum number of times to our server? Give the host name & count of connections from that host.**   
```{r q1, echo=TRUE}
RemoteHost.Sort <- data.frame(table(df_conn$RemoteHost))
r <- arrange(RemoteHost.Sort,desc(Freq))
head(r)
max(r$Freq)
```

**Which page that has been requested the maximum number of times from our server? Give the page name & count of the times the page was requested.**
```{r q2, echo=TRUE}
RequestLine.Sort <-data.frame(table(df_conn$RequestLine))
r <- arrange(RequestLine.Sort,desc(Freq))
r <- filter(r, !grepl("GET / HTTP/1.0",r$Var1))
head(r)
max(r$Freq)
```

**How many unique hosts have connected to our server? Give counts.**
```{r q3, echo=TRUE}
length(unique(df_conn$RemoteHost))  
```

**How many unique pages have been requested from our server? Give counts.**
```{r q4, echo=TRUE}
length(unique(df_conn$RequestLine)) 
```

**Which page has caused maximum data transfer from our server? Give page name & the data transfer for the page.**
```{r q5, echo=TRUE}
r <- summarise(group_by(dfPageSize, RequestLine), sum(Size))
r <- arrange(r,desc(`sum(Size)`))
r$RequestLine[1]
max(r$`sum(Size)`)
```

**Which host has caused maximum data transfer from our server? Give host name & the data transfer for the host.**
```{r q6, echo=TRUE}
r <- summarise(group_by(df_conn,RemoteHost ), sum(Size))
#r <- filter(r, !grepl(0,r$`sum(Size)`))
r <- arrange(r,desc(`sum(Size)`))
r$RemoteHost[1]
head(r)
```
**Which page has maximum download size from our server? Give page name & the size for the page.**
```{r q7, echo=TRUE}
s <- dfPageSize.Sort$RequestLine[1]
s
dfPageSize.Sort$Size[1] 
```
**What is the download count of the page that has maximum download size from our server? Give page name & download count**
```{r q8, echo=TRUE}
cnt_s <- filter(dfPageSize.Sort, grepl(s,dfPageSize.Sort$RequestLine))
s
nrow(cnt_s)
```
**Which page has minimum download size from our server? Give page name & the size for the page.**
```{r q9, echo=TRUE}
dfAscSize.Sort <- arrange(dfPageSize,Size)
m <- dfAscSize.Sort$RequestLine[1]
m
dfAscSize.Sort$Size[1]
```
**What is the download count of the page that minimum download size from our server? Give page name & the size for the page.**
```{r q10, echo=TRUE}
cnt_m <- filter(dfAscSize.Sort, grepl(m,dfAscSize.Sort$RequestLine))
nrow(cnt_m)
m
dfAscSize.Sort$Size[1]
```

